Estimation below the detection limit

### 2018-04-07 TEX/draft2.tex
updated notes

\documentclass[journal]{IEEEtran}

@@

\usepackage[T1]{fontenc}
\usepackage[latin9]{inputenc}
\usepackage{verbatim}
\usepackage{float}
\usepackage{amsthm}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{graphicx}
%\usepackage{multirow}
\usepackage{color}
\usepackage{url}

\usepackage{algorithm}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% Textclass specific LaTeX commands.
\numberwithin{equation}{section}
\numberwithin{figure}{section}
\theoremstyle{plain}
\newtheorem{thm}{\protect\theoremname}[section]
\theoremstyle{definition}
\newtheorem{defn}[thm]{\protect\definitionname}
\theoremstyle{remark}
\newtheorem{claim}[thm]{\protect\claimname}
\theoremstyle{plain}
\newtheorem{lem}[thm]{\protect\lemmaname}

\newtheorem*{lem*}{Lemma}
\theoremstyle{remark}
\newtheorem{rem}[thm]{\protect\remarkname}
\theoremstyle{plain}
\newtheorem{corollary}[thm]{\protect\corollaryname}
\theoremstyle{plain}
\newtheorem{proposition}[thm]{\protect\propositionname}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% User specified LaTeX commands.

%\usepackage{babel}
\providecommand{\claimname}{Claim}
\providecommand{\definitionname}{Definition}
\providecommand{\lemmaname}{Lemma}
\providecommand{\remarkname}{Remark}
\providecommand{\theoremname}{Theorem}
\providecommand{\corollaryname}{Corollary}
\providecommand{\propositionname}{Proposition}

\newcommand{\RL}{\mathbb{R}^L}
\newcommand{\RN}{\mathbb{R}^N}
\newcommand{\CL}{\mathbb{C}^L}
\newcommand{\CN}{\mathbb{C}^N}
\newcommand{\RW}{\mathbb{R}^W}
\newcommand{\RNN}{\mathbb{R}^{N\times N}}
\newcommand{\CNN}{\mathbb{C}^{N\times N}}
\newcommand{\inner}[1]{\left\langle {#1} \right\rangle}
\newcommand{\E}[1]{\mathbb{E}\left\{{#1} \right\}}
\newcommand{\order}[1]{\mathcal{O}\left({#1} \right)}
\newcommand{\xz}{x_{\textrm{zp}}}
\newcommand{\DFT}{\operatorname{DFT}}
\newcommand{\SNR}{\operatorname{SNR}}
\newcommand{\hx}{\hat{x}}

@@

%\begin{frontmatter}

\title{Multireference alignment meets blind deconvolution}
\author{Tamir Bendory}
\maketitle

@@

\begin{IEEEkeywords}
	Blind deconvolution, multireference alignment, bispectrum, cryo-EM
\end{IEEEkeywords}

\section{Introduction} \label{sec:introduction}

Blind deconvolution is a longstanding problem, arising in a variety of engineering and scientific applications, such as astronomy, communication, image deblurring, system identification and optics; see   \cite{jefferies1993restoration,tong1994blind,chan1998total,campisi2016blind,kundur1996blind,levin2011understanding,shalvi1990new,levin2009understanding,krishnan2011blind,ayers1988iterative,michaeli2014blind,lin2005relevant,abed1997blind}, just to name a few. In blind deconvolution, we aim at estimating  a signal from its convolution with an unknown kernel. 
Clearly, without additional information, the problem is ill-posed.

In this paper, we consider a particular case, in which one signal is of finite support, while the other is a binary sparse signal. Let us define the (zero-padded) signal 
$$\xz  = [x, \underbrace{0,0,\ldots,0}_{N-L \text{ zeros}}]\in\RN,\quad N\gg L,$$
and a binary signal $s\in\{0,1\}^N$. Then, our goal is to estimate $x$ from \begin{equation} \label{eq:blind_deconvolution}
y = \xz \ast s + \varepsilon,
\end{equation}
where $\varepsilon$ is an i.i.d.\ normal noise $\varepsilon$ with mean zero and variance $\sigma^2$.
As can be seen, the measurement $y\in\RN$ is composed of repetitions of an underlying signal $x\in\RL$, located at different positions. So, we can also write the model as  
\begin{equation}
y[n] = \sum_{i=1}^K \xz [n-n_i] + \varepsilon[n], 
\end{equation}
where the $n_i$'s are the non-zero values of $s$.

While most works on blind deconvolution tries to estimate both unknown signals (see Section~\ref{sec:blind_deconvolution}), in most cases, the goal is merely to estimate one of the signals.
For instance, in image deblurring, both the blurring kernel and the high-resolution image are unknown, but the main goal is only to sharpen the image. In~\eqref{eq:blind_deconvolution}, the signal $s$ is referred to as the latent or hidden variable of the problem. Our goal is to estimate  $x$ and we focus  on the low signal--to--noise ratio ($\SNR$) regime. As will be demonstrated, in this case, estimating $s$ is impossible, even if $\xz$ is known. However, we show, maybe surprisingly, that one can estimate $x$ accurately in some cases.  We support this statement by theoretical analysis and numerical  experiments.

Figure~\ref{fig:example} shows examples of the problem in different noise levels. In the example, the signal $x$ appears twice in $y$ so that $\| s \|^2 = 2$. The first row demonstrates a high $\SNR$ regime with $\sigma = 0.1$. As can be seen, in this regime  the problem is rather easy. In this case, one can easily the support of $s$ (namely, the repetition of $x$ in $y$) by simple detection algorithms. This is demonstrated by the correlation function between $\xz$ and $y$ that exhibits two unmistakable peaks in the support of $x$.  Once several copies of the signal were detected and cropped from $y$, then one can  improve the $\SNR$ by averaging. We are interest in the more challenging regime of high noise level, as exemplified in the bottom row ($\sigma  = 3$)
in which the signal is completely swamped in the noise, and one cannot detect the signal's appearances. In this case, one need to use the prior information that the signal appears many times across the measurement.

\begin{figure*}
	\begin{center}
	\includegraphics[scale = .5]{example}
	\end{center}
\caption{The left column presents the clean data and the measurement in different noise levels. In this example, a rectangular signal of length $L=21$ appears twice in a measurement of length $N=336$ with different noise levels. Note to the different scale of the y-axis. The right column presents the correlation between the rectangular signal and the measurement for the associated problems. }
\label{fig:example}
\end{figure*}

Our approach for this blind deconvolution problem is based on  tools that were developed for the multireference alignment (MRA) problem. MRA is the problem of estimating a signal from its circularly-shifted noisy measurements~\cite{bandeira2014multireference,bendory2017bispectrum} 
 In Section~\ref{sec:mra}, we survey recent works in this field. We use a recent suggestion to estimate the signal from feature that are invariant under cyclic-translation~\cite{bendory2017bispectrum}. 
 In essence, our method is based on \emph{approximate reduction} of the blind deconvolution problem into the MRA model. This is done be splitting $y$ into shorter windows and treat them as the MRA measurements. If each window contains exactly  one signal, then the problem reduces exactly the MRA problem. However, this is not the case. Some windows contain only noise, while others contain both noise and parts of a signal.  
 We analyze the estimation rate of this method and its deviation from the MRA model.

 A motivation for this work arises from the cryo--electron microscopy (EM) problem. Cryo--EM received a great deal of attention in structural biology in recent years as it allows visualizing molecules that were not stained in any way, showing them in their native environment. In Section~\ref{sec:cryoEM} we draw the connections between  blind deconvolution and the cryo--EM problem.

 The outline of this paper is as follows. In Section~\ref{sec:literature} we introduce background  on blind deconvolution (Section~\ref{sec:blind_deconvolution}) and MRA (Section~\ref{sec:mra}).  
 In Section~\ref{sec:cryoEM} we discuss the connection with the  cryo--EM problem.  In Section~\ref{sec:algorithm} we introduce the algorithm and analyze it in Section~\ref{sec:analysis}. Section~\ref{sec:experiments} is devoted for numerical experiments. Section~\ref{sec:conclusion} concludes the manuscript.

\section{background} \label{sec:literature}

In this section, we introduce the two main ingredients of this work, blind deconvolution and MRA. Then, we draw connections  with the cryo--EM problem.

\subsection{Blind deconvolution} \label{sec:blind_deconvolution}

Blind deconvolution is the problem of recovering a signal from its convolution with an unknown kernel $h$. That is, we aim to recover  $x$ from
\begin{equation}
y = x\ast h,
\end{equation}
where both $x$ and $h$ are unknown. In the Fourier domain, it takes the form of 
\begin{equation}
Fy = Fx \odot Fh,
\end{equation}
where $Fz$ stands for the Fourier transform of $z$ and $\odot$ for entry-wise product.
Clearly, without prior information of the signal, it is impossible to estimate the signal. For instance, even if we assume that the $\DFT$ of both signals are non-vanishing, then $Fy$ remains unchanged if we replace $(Fx)[k],(Fh)[k]$ by $a(Fx)[k],\frac{1}{a}(Fh)[k]$ for any $a\in\CN\backslash\{0\}$ and any $k$.

The blind deconvolution  has been analyzed recently with a various convex and non-convex algorithms. These works assume that the signal lies in a known  subspace spanned by vectors with random entries~\cite{ahmed2014blind,li2016rapid,ling2017blind}. A somewhat more related line of works assume sparsity in a random linear subspaces~\cite{lee2017blind,ling2015self,chi2016guaranteed}. 
The papers 
\cite{choudhary2014sparse,li2016identifiability,kech2017optimal,li2015unified} deals with the fundamental question of identifiability of the problems, but none of these paper consider our setup, namely, when the sparse signal is in the canonical basis.

Our problem is very different from all previous works in at least two ways. First, 
while we do assume that one of the signals is sparse,  our goal is to estimate the other signal. We do not claim that we can estimate the sparse signal. Second, we are mainly care of accurate estimation of $x$ in the low $\SNR$ regime. In this regime, it seems that estimating the sparse signal $s$ is impossible. This is contrast to previous works that focused on the high $\SNR$ regime.

We briefly mention that the deconvolution problem, namely, when the kernel is assumed to be known, has been analyzed throughly in recent years under sparsity constraints~\cite{bendory2016robust,bendory2017robust,bendory2016stable,boyer2017adapting,bernstein2017deconvolution,de2012exact,azais2015spike,duval2015exact,duval2015sparse}.  Nevertheless, the problem under consideration is essentially different  from two reasons. First, of course, we consider the blind setting when the kernel is unknown. Second, these works considered the recovery of sparse signal, with Gaussian-like kernel. In our problem, the convolution itself kernel is sparse, while we aim to recover a general signal. 
.

\subsection{Multireference alignment} \label{sec:mra}

Our approach for blind deconvolution is based on reducing the problem, approximately, to an MRA problem.
In MRA, we aim at estimating a signal $x\in\RL$ from its circularly shifted noisy measurements
\begin{equation} \label{eq:mra}
y_j = R_{r_j}x + \varepsilon_j, \quad j = 1,\ldots,N,
\end{equation}
where $R_{r}$ translates a signal by $r$ locations, namely, $(R_rx)[i] = x[(i-r)\mod L]$, and $\varepsilon$ is an i.i.d.\ normal noise with mean zero and variance $\sigma^2$. This problem finds applications in radar, image processing and structural biology~\cite{zwart2003fast,foroosh2002extension,diamond1992multiple}. The algorithmic and statistical characterizations of this problem have been analyzed thoroughly in the last couple of years, see~\cite{bendory2017bispectrum,bandeira2014multireference,bandeira2017optimal,perry2017sample,boumal2017heterogeneous,abbe2017multireference,abbe2017sample}.

In this work, we exploit the algorithmic framework proposed in~\cite{bendory2017bispectrum}. In order to estimate the signal, this paper suggests to use features of the signal that are invariant under cyclic translation.
Apparently, three moments are required. The first is mean of the signal, denoted by $\mu_x$, or, equivalently, it DC component $(Fx)[0]$. The second feature is the power spectrum of the signal $P_x[k] = \vert (Fx)[k]\vert^2$ for $k=0,1,\ldots L-1$. Since in general one cannot estimate  a signal from its power spectrum~\cite{bendory2017fourier}, one needs a third order feature. This third order feature, called bispectrum, is defined as~\cite{tukey1953}
\begin{equation}
B_x[k_1,k_2] = (Fx)[k_1]\overline{(Fx)[k_2]}(Fx)[k_2-k_1].
\end{equation}  
These features can be estimated directly from the measured data by
\begin{align} \label{eq:moment_estimation}
\frac{1}{N}\sum_{i=1}^N\mu_{y_j}  &\,\to \, \mu_x, \nonumber\\
\frac{1}{N}\sum_{i=1}^NP_{y_j}  &\,\to \, P_x + L\sigma^2, \\
\frac{1}{N}\sum_{i=1}^NB_{y_j}  &\,\to \, B_x + \mu_x\sigma^2L^2A,  \nonumber
\end{align} 
where $A$ is a known deterministic matrix. Therefore, by proper debiasing, one can get a reliable estimating of these features. For large enough $\sigma$, the variance of the estimators  is dominated by the bispectrum and thus goes as $\sigma^6/N$. This estimation rate is optimal under the assumption that distribution of translations is uniform~\cite{bandeira2017optimal}. If the distribution is non-uniform, then one can do better using only the first two moments, see~\cite{abbe2017multireference}. 
 Given enough measurements, one can obtain reliable estimation of the features, and then to estimate the underlying signal, up to cyclic translation, using a variety of algorithms~\cite{bendory2017bispectrum}. In this work, we use a simple non-convex least-squares estimator.... \TODO{Need to see what do we actually solve...}

\subsection{Connections with the cryo--EM problem} \label{sec:cryoEM}

This work is partially motivated by the imaging
technique called single particle cryo--EM, enabling the visualization of molecules at near atomic resolution~\cite{bartesaghi20152,sirohi20163}. Cryo--EM gained a lot of popularity in recent years in the structural biology community. As a result, Dubochet, Frank and Henderson were recently awarded the 
the Nobel prize for their contribution in the development of this technique~\cite{nobel}.

In cryo--EM, many samples of a molecule are frozen in a thin sheet of ice. 
An electron beam then passes through the ice and the samples  and recorded by an array of detectors. Cryo--EM contains several such images, which are
called \emph{micrographs}. Each micrograph  contains many two-dimensional (2D) tomographic projections of the samples. 
 The  cryo--EM problem is then to estimate the molecules from the micrographs. This process is illustrated by in Figures~\ref{fig:cryo-EM} and~\ref{fig:cryo-EM-problem}.

Compared to more traditional tomographic methods, like computerized tomography (CT), the cryo--EM problem arises two main challenges. First, the 3D orientation of each sample is unknown since the samples are rapidly frozen in the ice sheet. The experimentalist does not have control over these orientations. Second, the $\SNR$ in each projection is very low since the electron dosage must be restricted to limit  radiation damage to the molecules. In real cryo--EM data sets, there are some additional challenges, like the contrast transfer function of the microscope that corrupts the acquired image.

\begin{figure}
	\begin{center}
		\includegraphics[scale = .4]{cryoem-eps-converted-to}
	\end{center}
	\caption{A schematic draw of the micrograph generation. Courtesy of ?}
	\label{fig:cryo-EM}
\end{figure}

\begin{figure}
	\begin{center}
		\includegraphics[scale = .35]{cryoem-problem-eps-converted-to}		
	\end{center}
	\caption{The cryo--EM problem: estimating a molecule from the micrograph. Courtesy of ?}
	\label{fig:cryo-EM-problem}
\end{figure}

The first stage of a cryo--EM reconstruction algorithm is the so called \emph{particle picking} stage. In this stage, the projections are detected in the micrographs, and then cropped to a series of 2D images.  Many algorithms were proposed to this detection problem based on support vector machine classifiers, deep learning and template matching and more; see for instance \cite{frank1983automatic,scheres2015semi,arbelaez2011experimental,zhu2016deep}.
 The most popular techniques are based on the latter, where the templates can manually or automatically selected.  The underlying idea is that the cross-correlation  should be high in the presence of a signal in the template. However, this is not true anymore in the presence of high noise level, even if the signal itself is known, as demonstrated in Figure~\ref{fig:example}. An illuminating examples are given in~\cite{shatsky2009method,bandeira2015non}, where the famous picture of Albert Einstein is used to match pure noise and rotated images of the mathematician  Hermann Weyl.

After the particle picking, ideally, one gets a set of measurements of the form 
\begin{equation} \label{eq:cryo-em}
y_j = \mathcal{P}( g_j\circ x) + \varepsilon_j,\quad j=1,\ldots,N, 
\end{equation}
where $x$ is the three-dimensional (3D) molecule to be estimated,  $g_j\in SO(3)$ are unknown 3D rotations that act on the molecule and $\mathcal{P}$ is a linear tomographic projection. This model has been analyzed theoretically from different point--of-views~\cite{bandeira2015non,hadani2011representation}. The MRA model~\eqref{eq:mra} can be interpreted as a simplification of the model~\eqref{eq:cryo-em}, where the unknown cyclic-shifts correspond to the unknown rotations (but there is no analog for the projection operation).

In practice, particle picking algorithms are far from being optimal, mainly because the high noise level prevents reliable detection. One result is that the cropped images are usually not centered,  and therefore an unknown 2D translation should be added to the model~\eqref{eq:cryo-em}.
In addition, many samples in the micrograph are non detected or ignored.
After the particle picking stage, the molecule is estimated from the 2D images, where the most popular algorithms are based on expectation--maximization (see for instance~\cite{scheres2012relion}). Other techniques, such as the common-lines\cite{van1987angular,singer2011three} method or auto-correlation analysis~\cite{kam1980reconstruction,levin20173d}, can be used for ab initio modeling.

 Our model can be seen as a highly simplified cryo--EM model, where $y$ corresponds to the micrograph and the signal repeated $x$ to a  ``one-dimensional molecule" (the tomographic projection $\mathcal{P}$ does not appear in the model). We are interest in the question whether we can estimate the signal $x$, even if the $\SNR$ is so low that we cannot detect the locations of $x$, in $y$. In the perspective of the cryo--EM problem. the success to estimate the signal in the very low SNR regime may indicate that the molecule can be estimated directly from the cryo--EM micrographs, even if the SNR is so low that particle picking algorithm work poorly.

>>>>

\documentclass[english,11pt]{article}

@@

\usepackage[T1]{fontenc}
\usepackage[latin9]{inputenc}
\usepackage{verbatim}
\usepackage{float}
\usepackage{amsthm}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{graphicx}
%\usepackage{multirow}
\usepackage{color}
\usepackage{url}
\usepackage{caption}
\usepackage{subcaption}
\usepackage{mathtools} 
\usepackage[margin=1.2in]{geometry}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% Textclass specific LaTeX commands.
\numberwithin{equation}{section}
%\numberwithin{figure}{section}
\theoremstyle{plain}
\newtheorem{thm}{\protect\theoremname}[section]
\theoremstyle{definition}
\newtheorem{defn}[thm]{\protect\definitionname}
\theoremstyle{remark}
\newtheorem{claim}[thm]{\protect\claimname}
\theoremstyle{plain}
\newtheorem{lem}[thm]{\protect\lemmaname}

\newtheorem*{lem*}{Lemma}

\theoremstyle{remark}
\newtheorem{rem}[thm]{\protect\remarkname}
\theoremstyle{plain}
\newtheorem{corollary}[thm]{\protect\corollaryname}
\theoremstyle{plain}
\newtheorem{proposition}[thm]{\protect\propositionname}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% User specified LaTeX commands.
%\usepackage{slashbox}

\usepackage{babel}
\providecommand{\claimname}{Claim}
\providecommand{\definitionname}{Definition}
\providecommand{\lemmaname}{Lemma}
\providecommand{\remarkname}{Remark}
\providecommand{\theoremname}{Theorem}
\providecommand{\corollaryname}{Corollary}
\providecommand{\propositionname}{Proposition}

\newcommand{\reals}{\mathbb{R}}
\newcommand{\RL}{\mathbb{R}^L}
\newcommand{\CL}{\mathbb{C}^L}
\newcommand{\RN}{\mathbb{R}^N}
\newcommand{\RNN}{\mathbb{R}^{N\times N}}
\newcommand{\CNN}{\mathbb{C}^{N\times N}}
\newcommand{\inner}[1]{\left\langle {#1} \right\rangle}
\newcommand{\hx}{\hat{x}} 
\newcommand{\one}{\mathbf{1}} 
\newcommand{\SNR}{{\textsf{SNR}}}

@@

\title{Estimation below the detection limit}

\author{Tamir Bendory}
\maketitle

@@

\section{Introduction}

In this paper, we consider the problem of estimating a set of signals $x_1,\ldots,x_K$ from their multiple occurrences in unknown  locations in a data sequence $y$\TODO{This is the most important sentence of the paper, we need to polish it up}. The data may also contain background information---independent of the signals---modeled as noise.
For one-dimensional signals, the data sequence can be thought of as a 
long time series and the $K$ signals as repeated short events.  
This model appears in many applications, including spike sorting~\cite{lewicki1998review}, passive radar~\cite{gogineni2017passive} and system identification~\cite{ljung1998system}.
In the last part of this manuscript, we also propose to interpreted this estimation problem as a toy model for the problem of single particle reconstruction using cryo-electron microscopy (cryo--EM). A precise mathematical formulation of the model and the estimation problem is provided in Section~\ref{sec:model}.

Figure~\ref{fig:XC_example} shows a simple example of data sequences with  one signal ($K=1$) that appears twice.
The upper row presents the data without noise, in the left panel, and its cross-correlation with the signal itself in the right panel. The cross-correlation exhibits two peaks corresponding to the signal's locations in the data. 
%Even if we add a small amount of noise and considering more than one signal ($K>1$), estimating the signals remains a painless task. 
%In this scenario, standard detection and clustering algorithms can locate and cluster the copies of the signals that can be then averaged within each class.  
In  the middle row, the same data is shown, now contaminated with i.i.d.\ Gaussian noise with zero mean and standard deviation of $\sigma=0.7$. 
While it is harder to observe the signal occurrences in the left panel, the cross-correlation still produces fairly adequate estimates of the signal's locations. In practice one usually does not possess a precise template of the signal to be recovered, however, clever methods based on template matching, such as those used in structural biology~\cite{heimowitz2018apple} and radar~\cite{gogineni2017passive}, may work.
The bottom panels show the same data swamped in  
i.i.d.\ Gaussian noise with standard deviation  $\sigma=3$. In this  low signal--to--noise (\SNR) regime, 
detection of individual signal occurrences is impossible, 
even if the true signal is known. 
This phenomenon---explained in more detail in Section~\ref{sec:model}--raises the question: 
\medskip 
\\
\emph{Can we estimate the signals accurately without detecting their individual occurrences? }
 \\

In this work, we provide a positive answer to this question. For one signal ($K=1$), we prove that in the asymptotic regime in which the signal appears
infinitely many times in the data, and under a spacing condition, one can estimate the signal \emph{to any desired accuracy in any $\SNR$ level}. We show empirically that the same holds true for multiple signals if $K$ is not too large. \TODO{Do we want a sentence like: Based on  recent results on the multireference alignment problem, we conjecture that it remains true as long as $K \lesssim L/6$. ?}

Our framework is based on autocorrelation analysis.
In a nutshell, the method consists of two stages. First, we estimate a mixture (linear combination) of the low--order autocorrelation functions of the signals from the data. These quantities can be estimated, to any desired accuracy, if individual occurrences are separated by at least $L-1$ entries and  each signal appears ample number of times \TODO{?} in the data. There is no need to detect individual occurrences. 
Furthermore, we do not assume the knowledge of the number of signal occurrences $M$ and the noise level.
In the second stage, the signals are estimated from the mixed autocorrelations using a nonconvex least-squares (LS).
Section~\ref{sec:autocorrelation} elaborates on the technique and proves some of its properties, while Section~\ref{sec:numerics} provides numerical demonstrations.

Interestingly, expectation-maximization (EM)---a popular framework for similar estimation problems, such as Gaussian mixture models and multireference alignment---is intractable for this problem, even if the number of signal occurrences $M$ is known and $K=1$.
In particular, in each iteration, EM assigns probabilities to all possible combinations of positioning the current signal estimate in $M$ locations on $\mathbb{Z}_N$, resulting in $O(N^M)$ possibilities (even when excluding forbidden combinations due to the spacing constraint).

\begin{figure}
		\centering
		\includegraphics[scale=0.5]{XC_example}
		\caption{ The upper left panel shows a data sequence of length $N=80$ with one rectangular signal of length $L=11$ that appears twice. In the second and third rows of the left column, a Gaussian noise with mean zero and standard deviation $\sigma=0.5$ and $\sigma=3$ was added, respectively. The panels in the right column present the cross-correlation of the data with the true rectangular signal. Note to the different scales in different panels. }
    	\label{fig:XC_example}
\end{figure}

%
%This work is primely motivated by cryo-electron microscopy (cryo--EM), which is an innovative  single particle reconstruction technology. The acquired data in a cryo-EM experiment is contaminated with high noise levels. Therefore, any molecule reconstruction algorithm must take  the challenging $\SNR$ level into account.  In the last part of this manuscript, we draw connections with the estimation problem under consideration and the cryo--EM problem.

\section{Model and related literature}  \label{sec:model}

Let $x_1,\ldots,x_K\in\RL$ be the sought signals and let $y\in\RN$ be the data. 
A plain way to present the forward model is as a mixture of blind deconvolution problems between binary signals and the target signals
\begin{equation} \label{eq:model}
y = \sum_{i=1}^K x_i\ast s_i + \varepsilon,\quad \varepsilon\sim\mathcal{N}(0,\sigma^2 I).
\end{equation}
%We model the background information as i.i.d.\ Gaussian noise with zero mean and $\sigma^2$ variance. 
Each  $s_i\in\{0,1\}^N$ is associated with the position of the occurrences of the corresponding $x_i$. Let $\mathcal{S}_i$ denotes the set of  the nonzero
 values of $x_i$ and assume that the all $\mathcal{S}_i$'s are disjoint. The union of supports is denoted by $s = \sum_{i=1}^Ks_i$ and similarly $\mathcal{S} = \cup \mathcal{S}_i$. 
The cardinality of the sets are denoted by $\vert \mathcal{S}_i\vert = M_i$ and $\vert \mathcal{S}\vert :=M =  \sum_{i=1}^{K}M_i$.  Neither the $M_i$'s nor $M$ are assumed to be known.

In order to estimate the mixture of signals autocorrelations, we assume that the support of $s$ is not clustered. In particular, 
\begin{equation} \label{eq:spacing}
\text{For all }  i,j\in\mathcal{S}, \, i\neq j,  \quad  \vert i-j \vert\geq L-1.  
\end{equation}
The goal of the problme is to estimate $x_1,\ldots,x_K$ from $y$.

Blind deconvolution is a longstanding problem, arising in a variety of engineering and scientific applications, such as astronomy, communication, image deblurring, system identification and optics; see~\cite{jefferies1993restoration,shalvi1990new,ayers1988iterative,abed1997blind}, just to name a few. 
To make the problem well-posed, we must  assume some prior knowledge or  structure. 
In our case, the prior information is that $s$ is a binary signal that satisfies~\eqref{eq:spacing}. 
Other settings of blind deconvolution problems have been analyzed recently, see for instance~\cite{ahmed2014blind,li2016identifiability,li2016rapid,ling2015self,ling2017blind,chi2016guaranteed}
where the focus is on high $\SNR$ regimes.

An important feature of the problem under consideration is that while both $x_i$'s and $s_i$'s are unknown, the goal is merely to estimate the $x_i$'s. The  $s_i$'s  are referred to as \emph{nuisance  variables}. Indeed, in many blind deconvolution applications the sole purpose is to estimate one of the unknown signals. For instance, in image deblurring, both the blurring
kernel and the high-resolution image are unknown, but the primary goal is only
to sharpen the image.

If $x$ is known and $K=1$, then a sparse signal can be estimated by linear programming  in the high $\SNR$ regime, e.g.,~\cite{de2012exact,duval2015exact,bendory2016robust,bendory2017robust,bernstein2017deconvolution}. However, in the low $\SNR$ regime, estimating $s$ is impossible. To see that, suppose that an oracle provides us $M$ windows of length $W>L$, each contains one copy of $x$. That is to say, we get a series of windows, each one contains a signal at unknown location. 
Estimating the position of the known signal within each  window is an easier problem than detecting the support of $s$. 
However, even this problem is impossible in the low $\SNR$ regime~\cite{aguerrebere2016fundamental}. Therefore, we conclude that detecting the nonzero values of $s$ is impossible in low $\SNR$. As aforementioned, this work focuses on this regime and examines under what conditions we can estimate the signals themselves, despite the impossibility of detecting their individual occurrences.

For $K=1$, our problem can be also interpreted as a special case of the system identification problem. Similarly to~\eqref{eq:model}, the system identification forward model takes the form
\begin{eqnarray}
y = x\ast w + \varepsilon,  
\end{eqnarray} 
where $x$ is the unknown signal (``system''), $w$ is an unknown, random, input sequence and $\varepsilon$ is an additive noise.   
%The problem has also been studied in the case of a known input $w$~\cite{pillonetto2010new,dinuzzo2015kernels,bottegal2016robust}. 
The goal of this problem is to estimate $x$ (identify the system). The question of identifiability of $x$ under
this observation model is addressed for certain Gaussian and non-Gaussian $w$ in~\cite{benveniste1980robust,kormylo1983identifiability}.
In the special case where $w\in\{0,1\}^N$, satisfying the spacing requirement~\eqref{eq:spacing}, we obtain our
model in the  case of a single signal ($K = 1$).

Likelihood-based methods seek to maximize the likelihood function for $x$,
given the observed signal $y$. Solving this optimization exactly is typically
intractable, and so heuristic methods are used instead. One proposed heuristic
 is to use Markov Chain Monte Carlo (MCMC); in special
cases, including the case where $w$ is binary, EM has been used~\cite{cappe1999simulation}.
The EM method is based upon a certain ``forward-backward'' procedure
used in hidden Markov models, described in~\cite{rabiner1989tutorial}. \TODO{We need to explain}
Other problems solve for parameterized models of $x$; for instance, see~\cite{andrieu2001bayesian}. 
That same paper also considers multiple distinct signals $x$, as in our heterogeneity framework ($K>1$). Their proposed
solution is an MCMC algorithm designed for their specific parametrized problem.

Because likelihood methods are computationally expensive, methods based
on recovery from moments and cumulants, which are akin to our method, have
also been previously used for system identification. The question of identifiability
of $x$ from third-order statistics with non-Gaussian $w$ was considered in~\cite{lii1982deconvolution}. \TODO{how is it different from bispectrum inversion?}
Methods using the third- and fourth-order cumulants are described in~\cite{giannakis1989identification,tugnait1984identification}.

\section{Autocorrelation analysis}   \label{sec:autocorrelation}

Our method for estimating the signals is a two-stage technique. 
First, we use the autocorrelation functions of the data to estimate a mixture (i.e., linear combination) of the $K$ signal's autocorrelations. The mixed autocorrelation can be estimated to any accuracy, in any $\SNR$ level, if $M$ is large enough and the spacing condition~\eqref{eq:spacing} is met. Then, we  use a nonconvex LS  to estimate the signals from their mixed autocorrelations. 
In this section we elaborate on the autocorrelation functions and their estimations, while the precise recovery procedure, based on nonconvex optimization, will be discussed in detail in the next section.

\subsection{Aperiodic autocorrelation functions} \label{sec:aperiodic_ac}

For the purpose of this paper, we need the first three (aperiodic) autocorrelation functions. The first-order autocorelation is the mean of the signals. For  
$z\in\RL$ and $k\geq 2$, the higher order autocorrelations are defined by
\begin{equation} \label{eq:ac_general}
a_z^k[\ell_1,\ldots,\ell_{k-1}] =  \sum_{i=0}^{L-1-\max\{\ell_1,\ldots,\ell_{k-1}\}} z[i]z[i+\ell_1]\ldots z[i+\ell_{k-1}].
\end{equation}
Explicitly, the first three autocorrelations are
\begin{align} \label{eq:autocorrelations}
a_z^1 &= \sum_{i=0}^{L-1} z[i], \nonumber\\
a_z^2[\ell] &= \sum_{i=0}^{L-1-\ell} z[i]z[i+\ell], \nonumber\\
a_z^3[\ell_1,\ell_2] &= \sum_{i=0}^{L-1-\max\{\ell_1,\ell_2\}} z[i]z[i+\ell_1]z[i+\ell_2]. 
\end{align}
Note that the autocorrelation functions are symmetric so that $a_z^2[\ell] = a_z^2[-\ell]$ and $a_z^3[\ell_1,\ell_2] = a_z^3[-\ell_1,-\ell_2]$. 
Additionally, if the moments of the signal depend only on the difference between the indices (Toeplitz structure), then they are equivalent to the autocorrelation functions.

A one-dimensional signal is determined uniquely and stably by its third-order auto-correlation as proven in the following simple proposition.
\begin{proposition} \label{prop:uniqueness}
	Let $z\in\RL$ and suppose that $z[0]$ and $z[L-1]$ are nonzeros. Then:
	\begin{itemize}
		\item \textbf{Uniqueness:} 	 $z$  is determined uniquely from  $a_z^2$ and $a_z^3$.
        \item \textbf{Finite sensitivity:} 	Suppose that  $a_z^3[k,L-1]$ can be estimated up to  perturbation $\upsilon$ and let $\hat{z}[k]$ be an estimator of $z[k]$. If $\vert z[0]z[L-1]\vert \geq \delta>0$, then $\vert \hat{z}[k] - z[k]\vert\leq \frac{\vert \upsilon\vert }{\delta}$. 
	\end{itemize}
	\begin{proof}
		By assumption $a_z^2[L-1] = z[0]z[L-1]\neq 0$.
		Then, the uniqueness results, for all $k=0,\ldots L-1$,  follows from:
		\begin{equation*}
		a_z^3[k,L-1] = z[0]z[k]z[L-1].
		\end{equation*}
		If we measure $\tilde{a_z^3}[k,L-1] = a_z^3[k,L-1]+\upsilon$, then 
		\begin{equation*}
		\hat{z}[k] = \frac{\tilde{a_z^3}[k,L-1]}{a_z^2[L-1]} = z[k]+\frac{\upsilon}{a_z^2[L-1]} \quad \Rightarrow \quad \vert \hat{z}[k] - {z}[k]\vert \leq \frac{\vert\upsilon\vert}{\delta}.
		\end{equation*} 
	\end{proof}
\end{proposition}

A few remarks are in order. First, this result carries through to signals of any dimension.
Second, if the spacing condition~\eqref{eq:spacing} holds, then the length of the signal can be determined from the autocorrelations and 
therefore the assumption that the first and last entries are nonzero is met. In particular, if~\eqref{eq:spacing} holds for some spacing $W>L$, then $a_z^2[i]=0$ for all $i>L-1$.
Third, computing the $d$th autocorrelation amplifies the variance of the noise by a factor $d$ asymptotically. Therefore, if we can estimate $a_z^3$ up to small perturbation, it implies that we can estimate $a_z^2$ accurately as the proposition assumes. 
Finally, while Proposition~\ref{prop:uniqueness} shows only finite sensitivity, in practice LS estimator is robust to additive noise; see Section~\ref{sec:numerics}. This is compatible with results in related problems~\cite{bendory2017bispectrum,boumal2017heterogeneous}.

Considering the third-order autocorrelation is also a necessary condition to determine a signal from its autocorrelations. Indeed, the second-order autocorrelation of a one-dimensional signal does not determine a signals uniquely~\cite{beinert2015ambiguities,bendory2017fourier}. On the other hand, for dimensions greater than one, almost all signals are determined uniquely, up to sign (phase in the complex case) and reflection through the origin (with conjugation in the complex case)~\cite{hayes1982reconstruction,hayes1982reducible}. The sign ambiguity can be resolved by the mean of the signal if it is not zero. However,  in order to determine the reflection symmetry, one needs to use additional information.
